---
title: 'STAT 413/613 HW 2: stringr and lubridate'
author: "Jiarong Li"
date: "`r Sys.Date()`"
output:
  html_document:
    number_sections: yes
  word_document: default
editor_options:
  chunk_output_type: inline
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.align  = "center",
                      fig.height = 3, 
                      fig.width  = 5)
```

# Instructions {-}

Rename the starter file under the analysis directory as `hw_01_yourname.Rmd` and use it for your solutions.   
1. Modify the "author" field in the YAML header.  
2. Stage and Commit R Markdown and HTML files (no PDF files).   
3. **Push both .Rmd and HTML files to GitHub**.   
- Make sure you have knitted to HTML prior to staging, committing, and pushing your final submission.  
4. **Commit each time you answer a part of question, e.g. 1.1**   
5. **Push to GitHub after each major question**, e.g., Scrabble and Civil War Battles   
- **Committing and Pushing are graded elements for this homework.**  
6. When complete, submit a response in Canvas  
    
- Only include necessary code to answer the questions.
- Most of the functions you use should be from the tidyverse. Too much base R will result in point deductions.
- Use Pull requests and or email to ask me any questions. If you email, please ensure your most recent code is pushed to GitHub.
- Learning objectives:
    - Manipulate dates and times with lubridate.


# Scrabble Words

For this exercise, we are using the [Collins Scrabble Words](https://en.wikipedia.org/wiki/Collins_Scrabble_Words), 
which is most commonly used outside of the United States. The dictionary most often used in the United States is the 
[Tournament Word List](https://en.wikipedia.org/wiki/Official_Tournament_and_Club_Word_List).

**WARNING**: Do not try `str_view()` or `str_view_all()` on these data.It will stall your computer.

1. Use a readr function to load the 2015 list of Collins Scrabble Words into R from your data folder or from 
   <https://data-science-master.github.io/lectures/data/words.txt>  
   + (note: "NA" is an official Scrabble word).
```{r 1}
library(tidyverse)
library(lubridate)
data <- read.table("https://data-science-master.github.io/lectures/data/words.txt", header = F, sep = ",")
head(data)
```
2. What are the six longest words that have the most "X"'s in them?
```{r 2}
str_length(data)
data %>%
  mutate(length = str_length(V1))%>%
  arrange(desc(length)) %>%
  filter(str_detect(V1, "X") == "TRUE") %>%
  mutate(containX = str_count(V1,"X"))%>%
  arrange(desc(containX)) -> data2
head(data2)
# The six longest are: COEXECUTRIXES, EXTRATEXTUAL, COEXECUTRIX, EXECUTRIXES, SAXITOXINS, XANTHOXYLS.
```
3. How many words have an identical first and second half of the word? If a word has an odd number of letters, exclude the middle character.  
  - MURMUR counts because MUR is both the first and second half.
  - JIGAJIG counts because the middle A is excluded so JIG is both the first and second half. 
  - Save the results to a variable.
```{r 3}
head(data)
data %>%
  mutate(halfword = (substr(V1, 1, floor(str_length(V1)/2)) == substr(V1, ceiling(str_length(V1)/2)+1, str_length(V1)))) %>%
  filter(halfword == "TRUE") -> data3
nrow(data3)
# There are 254 variables.
```
4. Use the results from 3 to find the longest word with an identical first and second half
   of the word?
```{r 4}
data3 %>%
  mutate(wordlength = str_length(V1))%>%
  arrange(desc(wordlength)) -> data4
head(data4)
# CHIQUICHIQUI is the longest.
```

# Civil War Battles

The data in "civil_war_theater.csv" contains a information on American Civil War battles, taken from [Wikipedia](https://en.wikipedia.org/wiki/List_of_American_Civil_War_battles).  

Variables include:   

- `Battle`: The name of the battle.
- `Date`: The date(s) of the battle in different formats depending upon the length of the battle. 
  + If it took place on one day, the format is "month day, year". 
  + If it took place over multiple days, the format is "month day_start-day_end, year". 
  + If it took place over multiple days and months, the format is "month_start day_start - month_end day_end, year". 
  + If it took place over multiple days,months, and years, the format is "month_start day_start, year_start - month_end day_end, year_end".
- `State`: The state where the battle took place. Annotations (e.g.     describing that the state was a territory at the time) are in parentheses.
- `CWSAC`: A rating of the military significance of the battle by the Civil War Sites Advisory Commission. `A` = Decisive, `B` = Major, `C` = Formative, `D` = Limited.
- `Outcome`: Usually `"Confederate victory"`, `"Union victory"`, or `"Inconclusive"`, followed by notes.
- `Theater`: An attempt to to identify which theater of war is most associated with the battle
    
1. Use a readr function and relative paths to load the data into R.
```{r 5}
cw <- read.csv("~/Documents/Master/Semester-3/STAT-613/hw-02-jiarong0829/data/civil_war_theater.csv", header = T)
head(cw)
```
The next several questions will help you take the dates from all the different formats and create a consistent set of start date and end date variables in the data frame. We will start by calculating how many years, and months are in each battle.

2. Add a variable to the data frame with the number of years for each battle.  

- Create a character variable as follows. This can be used as a pattern in a regular expression.  
   
```{r 6}
year_regex <- stringr::str_c(1861:1865, collapse = "|")
year_regex

```
    
- Use `year_regex` to now count the number of years in each battle, add this to the data frame,  and save the data frame.
```{r 7}   
cw %>%
  mutate(years.numbers = str_count(Date, year_regex)) -> cw1
```
3. Add a variable to the data frame with the number of months for each battle.  
- Consider R's built-in vector of month names: `month.name`.

```{r 8}
month.name <- stringr::str_c(c("January", "February", "March", "April", "May", "June","July", "August","September", "October","November", "December"), collapse = "|")
```
- Use `month.name` to count the number of month names in the `Date` variable in each battle. 
- Add this to the data frame. (You might need to do something similar to what we did in part 2).
```{r 9}
cw1 %>%
  mutate(month.count = str_count(Date, month.name)) -> cw2
```
    
4. Add a variable to the data frame that is `TRUE` if `Date` spans multiple days and is `FALSE` otherwise. Spanning multiple months and/or years also counts as `TRUE`.
```{r 10}
cw2 %>%
  mutate(span = str_detect(Date, "-")) -> cw3
```
   
5. Make four new data frames by filtering the data based on the length of the battles:  
   + a data frame with the data for only those battles spanning just one day, 
   + a data frame with the data for only those battles spanning multiple days in just one month, 
   + a data frame with the data for only those battles spanning multiple months but not multiple years, and,
   + a data frame with the data for only those battles spanning multiple years.
```{r 11}
cw3 %>%
  filter(span == 'FALSE') -> cw4
cw3 %>%
  filter(span == 'TRUE', month.count == 1) -> cw5
cw3 %>%
  filter(month.count > 1,years.numbers ==1) -> cw6
cw3 %>%
  filter(years.numbers > 1) -> cw7
```
6. For each of the four new data frames,  
  - Add two new variables:  
    + `Start` should contain the start-date of each battle. 
    + `End` should contain the end-date of each battle.
    + Hint: look at help for `separate()`
    + Make sure these are `Date` class objects. 
  - Remove the `Date` variable from  each data frame.
  - Save the data frames with the new variables
```{r 12}
cw4 %>%
  mutate(Start = mdy(Date), End = mdy(Date)) %>%
  select(Battle, State, CWSAC, Theater, Outcome, Start, End) -> cw8

cw5 <- separate(cw5, Date, into = c("Month","day", "year"), sep = " ")
cw5 <- separate(cw5, day, into = c("daystart","dayend"), sep = "-")
cw5 %>%
  mutate(Start = mdy(str_c(Month, daystart, ",", year))) %>%
  mutate(End = mdy(str_c(Month, dayend, year))) %>%
  select(Battle, State, CWSAC, Theater, Outcome, Start, End) -> cw9

cw6 <- separate(cw6, Date, into = c("Month","Year"), sep = ",")
cw6 <- separate(cw6, Month, into = c("Startday", "Endday"), sep = "-")
cw6 %>%
  mutate(Start = mdy(str_c(Startday,",",Year))) %>%
  mutate(End = mdy(str_c(Endday,",",Year))) %>%
  select(Battle, State, CWSAC, Theater, Outcome, Start, End) -> cw10

cw7 <- separate(cw7, Date, into = c("S","E"), sep = "-")
cw7 %>%
  mutate(Start = mdy(S))%>%
  mutate(End = mdy(E)) %>%
  select(Battle, State, CWSAC, Theater, Outcome, Start, End) -> cw11
```
7. Create a new data frame with all the battles and the Start and End  dates by binding the rows of the four data frames as updated in part 6 
```{r 13}
cw12 <- rbind(cw8, cw9, cw10, cw11)
head(cw12)
``` 
8. Calculate the number of days each battle spanned.  
- What's the longest battle of the war? 
- How long did it last?
```{r 14}
cw12 %>%
  mutate(daylength = End - Start) %>%
  arrange(desc(daylength)) -> cw13
# the longest battle of the war is Siege of Port Hudson which last 49 days.
```

9. Is there an association between the CWSAC significance of a battle and its duration?  
- Create an appropriate plot. 
- Interpret the plot in one sentence to answer the question.
- Extra Credit: Test for a linear relationship using `lm()` and interpret the results in one sentence based on the $p$-value and adjusted R-squared.
```{r, echo=TRUE, eval=TRUE}
ggplot(data = cw13)+
  geom_boxplot(mapping = aes(x = CWSAC, y = as.numeric(daylength))) +
  scale_y_log10()
# According to the boxplot, we can see C has the lowest duration. However, the duration of other three types are same. A has the longer duration among four CWSAC.

cw.lm <- lm(as.numeric(daylength) ~ CWSAC, data = cw13)
summary(cw.lm)
# CWSAC is significant with lowever p-value and it can explain 5.48% of the data in this dataset.
```

10. Extra Credit: Did the [theaters of war](https://en.wikipedia.org/wiki/Theater_(warfare)) shift during the American Civil War?  
- Reproduce this plot in R:
- Interpret the plot in one sentence.
- Hints: 
  + Filter out states with two or fewer battles. 
  + Use regex to clean up the state names and then convert to factors
  + Use forcats to reorder the states by the start date. 
  + Use `coord_flip()` to make horizontal boxplots.
```{r, echo=TRUE, eval=TRUE}

```

       




















